Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-15 GWLDCII.tex
little things throughout

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is to enumerate the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid.

@@

We also impose the convention that all vertices should always be interpreted in terms of the cyclic order on $V$.  So, for any $i\in V$ we will use $i+1$ to denote the successor of $i$ in $V$.   In particular, if $V = [n]$ then $(n,n+3)$ means the propagator connecting edges $3$ and $n$. This viewpoint is beneficial for us since many of our proofs rely on restricting our attention to a region of the diagram bounded by a given propagator, or to short propagators such as $(i,i+2)$, and these behave the same regardless of whether we cross the ``first'' vertex or not. However, we emphasize that this is different to the standard convention in the physics literature, where propagators are written $(i,j)$ with $1 \leq i<j \leq n$.

We note that in this paper, we allow for the propagator set $\cP$ to be empty. This is for mathematical convenience, even though these diagrams are do not correspond to tree level scattering amplitudes.

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is comprised of two main components: an interaction matrix $C(W)$, and a volume form $\Omega(W)$. These are described next.

@@

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its non-zero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible WLD to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{figure}[h]
\[
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 1.43*\step:\zero+6.6*\step:\radius)
  \newnode{1}{}
  \newnode{3.55}{}
  \newnode{4.45}{}
  \newnode{7}{}
    \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.3}{6.7}{20}
  \newpropbend{2.9}{5.1}{45}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.43) {\footnotesize \em $\leq 6$ vertices in the \\ \footnotesize \em highlighted region};
\end{scope}
\end{tikzpicture}
\qquad \qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 3.67*\step:\zero+4.43*\step:\radius)
  \newnode{1}{i}
  \newnode[below]{2}{i+1\ \ \ }
  \newnode[below]{3}{i+2}
  \newnode[below]{5}{j+1}
  \newnode[below]{6}{\quad j+2}
  \newnode[right]{7}{j+3}
      \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.35}{3.6}{55}
  \newpropbend{4.4}{6.6}{55}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.5) {\footnotesize \em $\leq 2$ vertices in the highlighted \\ \footnotesize \em region, and  no other propagators \\ \em \footnotesize end in the highlighted region};
\end{scope}
\end{tikzpicture}
\]
\caption{The two configurations described in Lemma~\ref{lem sian}; every weakly admissible Wilson loop diagram with no non-supporting vertices admits at least one example of at least one of these configurations.}\label{fig: small configs}
\end{figure}

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible WLD with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs:
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 on one side of it and nothing else on that side.\label{item big and 2}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
See Figure~\ref{fig: small configs} for an illustration of the two configurations.
\end{lem}

@@

Let $p_{r+1} = (j_{r+1}, j_{r+1}+2)$ be a propagator of length 2 in $W_r$; if it forms part of configuration \ref{item big and 2} or \ref{item pair of 2s}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

@@

\begin{rmk}
In the case that all vertices of an admissible WLD $W$ support at least two propagators, then Lemma~\ref{lem sian} substantially simplifies.  By Remark~\ref{rem:props of length 2 and 3}, $W$ has no propagators of length $3$.  Configuration \ref{item big and 2} necessarily entails vertices with support $1$ as does configuration \ref{item pair of 2s} unless $j=i+2$.  So in the case that $W$ has all vertices with support at least two then $W$ must contain a pair of propagators of length $2$  with the property that the first propagator is $(i, i+2)$, the second is $(i+2, i+4)$ and no other propagator ends on the edge $i+2$.
\end{rmk}

@@

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators and $n$ vertices, then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ distinct vertices to each $I_i$ in $n$ steps.
\end{cor} 
\begin{proof}
Fix $i \in [n]$. It follows from the proof of Lemma~\ref{lem no fourth vertex} that in applying Algorithm~\ref{alg:put GN on WLD} to construct the set $I_i$, the algorithm can never reach the fourth vertex of a propagator's support (ordered with respect to $<_i$) without having already assigned that propagator. Therefore if the algorithm starts at vertex $i$, all propagators must have been eliminated by the time it gets to vertex $i-1$, ensuring that the algorithm is completed in less than one full circumnavigation of the diagram and that $I_i$ contains exactly $k$ distinct elements.
\end{proof}

@@

    \caption{Four cases for admissible WLDs with no non-supporting vertices. The gray half-propagators illustrate where propagators may occur, but are not required to exist.}\label{fig 3 cases}

@@

\caption{Example WLD for illustrating Algorithm~\ref{alg WLD to denom via GN} and bijections between propagators and vertices for each Grassmann necklace element.}\label{fig R eg}

>>>>

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is enumerating the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid.

@@

We also impose the convention that all vertices should always be interpreted in terms of the cyclic order on $V$.  So, for any $i\in V$ we will use $i+1$ to denote the successor of $i$ in $V$.   In particular, if $V = [n]$ then $(n,n+3)$ means the propagator connecting edges $3$ and $n$. This viewpoint is beneficial for us since many of our proofs rely on restricting our attention to a region of the diagram bounded by a given propagator, or to short propagators such as $(i,i+2)$, and these behave the same way regardless of whether we cross the ``first'' vertex or not. However, we emphasize that this is different to the standard convention in the physics literature, where propagators are written $(i,j)$ with $1 \leq i<j \leq n$.

We note that in this paper, we allow for the propagator set $\cP$ to be empty. This is for mathematical convenience, even though these diagrams do not correspond to tree level scattering amplitudes.

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is built from two main components: an interaction matrix $C(W)$, and a volume form $\Omega(W)$. These are described next.

@@

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its non-zero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible Wilson loop diagram to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{figure}[h]
\[
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 1.43*\step:\zero+6.6*\step:\radius)
  \newnode{1}{}
  \newnode{3.55}{}
  \newnode{4.45}{}
  \newnode{7}{}
    \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.3}{6.7}{20}
  \newpropbend{2.9}{5.1}{45}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.43) {\footnotesize \em $\leq 6$ vertices in the highlighted \\ \footnotesize \em region and no other propagators\\ \footnotesize \em in the highlighted region};
\end{scope}
\end{tikzpicture}
\qquad \qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 3.67*\step:\zero+4.43*\step:\radius)
  \newnode{1}{i}
  \newnode[below]{2}{i+1\ \ \ }
  \newnode[below]{3}{i+2}
  \newnode[below]{5}{j+1}
  \newnode[below]{6}{\quad j+2}
  \newnode[right]{7}{j+3}
      \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.35}{3.6}{55}
  \newpropbend{4.4}{6.6}{55}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.5) {\footnotesize \em $\leq 2$ vertices in the highlighted \\ \footnotesize \em region, and  no other propagators \\ \em \footnotesize end in the highlighted region};
\end{scope}
\end{tikzpicture}
\]
\caption{The two configurations described in Lemma~\ref{lem sian}; every weakly admissible Wilson loop diagram with no non-supporting vertices admits at least one example of at least one of these configurations.}\label{fig: small configs}
\end{figure}

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible Wilson loop diagram with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs:
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 on one side of it and nothing else on that side.\label{item big and 2}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
See Figure~\ref{fig: small configs} for an illustration of the two configurations.
\end{lem}

@@

Let $p_{r+1} = (j_{r+1}, j_{r+1}+2)$ be a propagator of length 2 in $W_r$.  If it forms part of configuration \ref{item big and 2} or \ref{item pair of 2s}
%the configuration \eqref{eq:2 prop config}
then we are done, so assume otherwise.

@@

\begin{rmk}
In the case that all vertices of an admissible Wilson loop diagram $W$ support at least two propagators, then Lemma~\ref{lem sian} substantially simplifies.  By Remark~\ref{rem:props of length 2 and 3}, $W$ has no propagators of length $3$.  Configuration \ref{item big and 2} necessarily entails vertices with support $1$ as does configuration \ref{item pair of 2s} unless $j=i+2$.  So in the case that $W$ has all vertices with support at least two then $W$ must contain a pair of propagators of length $2$  with the property that the first propagator is $(i, i+2)$, the second is $(i+2, i+4)$ and no other propagator ends on the edge $i+2$.
\end{rmk}

@@

\begin{cor}\label{GN alg well defined}
If $W$ is an admissible Wilson loop diagram with $k$ propagators and $n$ vertices, then Algorithm~\ref{alg:put GN on WLD} assigns exactly $k$ distinct vertices to each $I_i$ in at most $n$ steps.
\end{cor} 
\begin{proof}
Fix $i \in [n]$. It follows from the proof of Lemma~\ref{lem no fourth vertex} that in applying Algorithm~\ref{alg:put GN on WLD} to construct the set $I_i$, the algorithm can never reach the fourth vertex of a propagator's support (ordered with respect to $<_i$) without having already assigned that propagator. Therefore if the algorithm starts at vertex $i$, all propagators must have been eliminated by the time it gets to vertex $i-1$, ensuring that the algorithm is completed in at most one full circumnavigation of the diagram and that $I_i$ contains exactly $k$ distinct elements.
\end{proof}

@@

    \caption{Four cases for admissible Wilson loop diagrams with no non-supporting vertices. The gray half-propagators illustrate where propagators may occur, but are not required to exist.}\label{fig 3 cases}

@@

\caption{Example Wilson loop diagram for illustrating Algorithm~\ref{alg WLD to denom via GN} and bijections between propagators and vertices for each Grassmann necklace element.}\label{fig R eg}
